\chapter{Critical Points in Two Variable Functions}
\label{chap:critical-points-in-two-variable-functions}
\lecture{11}{4 Jun. 11:39}{Eleventh Lecture}

A function \(z=f(x,y)\) is a function with two variables. It may have
\emph{Local or Global \textbf{Minima} or \textbf{Maxima} or
\textbf{Saddle Point}}.
One can also check the concavity of the surface.

\section{Critical Points}
\label{sec:critical-points}
A critical point is a point \((a,b,f(a,b))\) where \(f_x=f_y=0\) or \(f_x\) or \(f_y\) is
undefined.

\section{Saddle Points}
\label{sec:saddle-points}
If a point \((a,b,f(a,b))\) is a \href{sec:critical-points}{stationary point}
and if,
\[\exists (x,y) \text{ s.t. } f(x,y)>f(a,b) \text{ somewhere } \land
f(x,y)<f(a,b) \text{ elsewhere }\]
Then, \((a,b,f(a,b))\) is a saddle point.

\section{Hessin/Discriminant of \(f\): \(D_f\)}
\label{sec:hessin-discriminant-of-f-d_f}
\[
  D_f = 
  \left|
  \begin{matrix}
    f_{xx} & f_{xy} \\
    f_{yx} & f_{yy}
  \end{matrix}
  \right| = f_{xx}\cdot f_{yy} - {f_{xy}}^2
\]
Here, at \((a,b)\),
\[ f_{xy}=f_{yx} \] 
Since f is continuos at
\((a,b)\). See \href{def:second-order-partial-derivative}{Clairaut's Theorem}.

\section{Confirming Extremes or Saddle}
\label{sec:confirming-extremes-or-saddle}
Check at \((x,y)=(a,b)\).

\begin{align}
  \textbf{Local Maxima }&: f_{xx}<0 \text{ and } D_f
  >0)\label{eqn:local-maxima-2vf}\\
  \textbf{Local Minima }&: f_{xx}>0 \text{ and } D_f<0)
  \label{eqn:local-minima-2vf}\\
  \textbf{Saddle }&: f_{xx}=0 \text{ and } D_f<0) \label{eqn:saddle-2vf}\\
  \textbf{Inconclusive }&:  D_f=0) \label{eqn:inconclusive-2vf}
\end{align}

\section{Finding Absolute Extremas}
\label{sec:finding-absolute-extremas}
A boundry will be given. To find the extremas,
\begin{enumerate}
  \item find all the critical points inside the boundry.
  \item find all the critical points in the boundry( single variable function)
  \item Compare the values of the function for all candidates
  \item List the absolute extremas(points of maximum and minimum value of the
    function).
\end{enumerate}
