\section{Introduction and Definition of AI}
\label{sec:intro-def-ai}
There are two main approaches to defining artificial intelligence (AI): one based on human behavior or thinking, and the other based on rationality.

\subsection*{Think Like Humans}
\label{ssec:think-like-humans-def-ai}
\begin{quote}
The automation of activities that we associate with human thinking, activities such as decision-making, problem-solving, learning \ldots''\\ \hspace*{\fill}--- Bellman, 1978 \end{quote} \begin{quote} AI is a system that can think like a human. Thinking humanly means cognitive modelling. We need to get inside the actual workings of human minds.''\
\hspace*{\fill}--- Russell & Norvig, 2016
\end{quote}

\subsection*{Act Like Humans}
\label{ssec:act-like-humans-def-ai}
\begin{quote}
The art of creating machines that perform functions that require intelligence when performed by people.''\\ \hspace*{\fill}--- Kurzweil, 1990 \end{quote} \begin{quote} The exciting new effort to make computers think \ldots machines with minds, in the full and literal sense.''\
\hspace*{\fill}--- Haugeland, 1985
\end{quote}
\begin{quote}
``We probably already know about the Turing test, where one of the humans will become the questioner asking the same question to both system and a real human. If the system reacts the same way as humans do, then we can consider that system is an AI.''\
\hspace*{\fill}--- Russell & Norvig, 2016
\end{quote}

\subsection*{Think Rationally}
\label{ssec:think-rational-def-ai}
\begin{quote}
The study of mental faculties through the use of computational models.''\\ \hspace*{\fill}--- Charniak \& McDermott, 1985 \end{quote} \begin{quote} In here we use the laws of thought approach by using logics. Do you know Aristotle’s syllogisms? They provide patterns for argument structures that always produce correct conclusions when given correct premises.''\
\hspace*{\fill}--- Russell & Norvig, 2016
\end{quote}

\subsection*{Act Rationally}
\label{ssec:act-rationl-def-ai}
\begin{quote}
Acting rationally means acting so as to achieve one's goals, given one's beliefs. An agent is just something that perceives and acts. In this approach, AI is viewed as the study and construction of rational agents.''\\ \hspace*{\fill}--- Russell \& Norvig, 2010 \end{quote} \begin{quote} A system is rational if it does the right thing given what it knows. So, in other words, if it doesn’t have enough information it can’t act rationally.''\
\hspace*{\fill}--- Russell & Norvig, 2016
\end{quote}
